---
title: "Homework 8"
author: "Emma Privett"
date: "2025-03-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Running the sample code

```{r, eval=TRUE,echo=TRUE}
#open libraries
library(ggplot2) # for graphics
library(MASS) # for maximum likelihood estimation

# quick and dirty, a truncated normal distribution to work on the solution set

#read in data vector
z <- rnorm(n=3000,mean=0.2)
z <- data.frame(1:3000,z)
names(z) <- list("ID","myVar")
z <- z[z$myVar>0,]
str(z)
summary(z$myVar)

#histogram
p1 <- ggplot(data=z, aes(x=myVar, y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2)
print(p1)

#empirical density curve
p1 <-  p1 +  geom_density(linetype="dotted",size=0.75)
print(p1)

#maximum likelihood parameters
normPars <- fitdistr(z$myVar,"normal")
print(normPars)
str(normPars)
normPars$estimate["mean"] # note structure of getting a named attribute


#normal
meanML <- normPars$estimate["mean"]
sdML <- normPars$estimate["sd"]

xval <- seq(0,max(z$myVar),len=length(z$myVar))

stat <- stat_function(aes(x = xval, y = ..y..), fun = dnorm, colour="red", n = length(z$myVar), args = list(mean = meanML, sd = sdML))
p1 + stat

#exponential
expoPars <- fitdistr(z$myVar,"exponential")
rateML <- expoPars$estimate["rate"]

stat2 <- stat_function(aes(x = xval, y = ..y..), fun = dexp, colour="blue", n = length(z$myVar), args = list(rate=rateML))
p1 + stat + stat2

#uniform
stat3 <- stat_function(aes(x = xval, y = ..y..), fun = dunif, colour="darkgreen", n = length(z$myVar), args = list(min=min(z$myVar), max=max(z$myVar)))
p1 + stat + stat2 + stat3

#gamma
gammaPars <- fitdistr(z$myVar,"gamma")
shapeML <- gammaPars$estimate["shape"]
rateML <- gammaPars$estimate["rate"]

stat4 <- stat_function(aes(x = xval, y = ..y..), fun = dgamma, colour="brown", n = length(z$myVar), args = list(shape=shapeML, rate=rateML))
p1 + stat + stat2 + stat3 + stat4


#beta
pSpecial <- ggplot(data=z, aes(x=myVar/(max(myVar + 0.1)), y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) +
  xlim(c(0,1)) +
  geom_density(size=0.75,linetype="dotted")

betaPars <- fitdistr(x=z$myVar/max(z$myVar + 0.1),start=list(shape1=1,shape2=2),"beta")
shape1ML <- betaPars$estimate["shape1"]
shape2ML <- betaPars$estimate["shape2"]

statSpecial <- stat_function(aes(x = xval, y = ..y..), fun = dbeta, colour="orchid", n = length(z$myVar), args = list(shape1=shape1ML,shape2=shape2ML))
pSpecial + statSpecial
```

## 2. Now it's my turn!

```{r, eval=TRUE,echo=TRUE}
z <- read.csv("kbsampledata.csv", header=TRUE,sep=",")
names(z) <- list("ID","myVar")
z <- z[z$myVar>0,]
str(z)
summary(z)

#histogram
p1 <- ggplot(data=z, aes(x=myVar, y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2)
print(p1)

#empirical density curve
p1 <-  p1 +  geom_density(linetype="dotted",size=0.75)
print(p1)

#maximum likelihood parameters
normPars <- fitdistr(z$myVar,"normal")
print(normPars)
str(normPars)
normPars$estimate["mean"] # note structure of getting a named attribute


#normal
meanML <- normPars$estimate["mean"]
sdML <- normPars$estimate["sd"]

xval <- seq(0,max(z$myVar),len=length(z$myVar))

stat <- stat_function(aes(x = xval, y = ..y..), fun = dnorm, colour="red", n = length(z$myVar), args = list(mean = meanML, sd = sdML))
p1 + stat

#exponential
expoPars <- fitdistr(z$myVar,"exponential")
rateML <- expoPars$estimate["rate"]

stat2 <- stat_function(aes(x = xval, y = ..y..), fun = dexp, colour="blue", n = length(z$myVar), args = list(rate=rateML))
p1 + stat + stat2

#uniform
stat3 <- stat_function(aes(x = xval, y = ..y..), fun = dunif, colour="darkgreen", n = length(z$myVar), args = list(min=min(z$myVar), max=max(z$myVar)))
p1 + stat + stat2 + stat3

#gamma
gammaPars <- fitdistr(z$myVar,"gamma")
shapeML <- gammaPars$estimate["shape"]
rateML <- gammaPars$estimate["rate"]

stat4 <- stat_function(aes(x = xval, y = ..y..), fun = dgamma, colour="brown", n = length(z$myVar), args = list(shape=shapeML, rate=rateML))
p1 + stat + stat2 + stat3 + stat4

#beta
pSpecial <- ggplot(data=z, aes(x=myVar/(max(myVar + 0.1)), y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) +
  xlim(c(0,1)) +
  geom_density(size=0.75,linetype="dotted")

betaPars <- fitdistr(x=z$myVar/max(z$myVar + 0.1),start=list(shape1=1,shape2=2),"beta")
shape1ML <- betaPars$estimate["shape1"]
shape2ML <- betaPars$estimate["shape2"]

statSpecial <- stat_function(aes(x = xval, y = ..y..), fun = dbeta, colour="orchid", n = length(z$myVar), args = list(shape1=shape1ML,shape2=shape2ML))
pSpecial + statSpecial
```

## 3. Which distribution fits the best?

Gamma!

## 4. Simulate a new data set

Here are my maximum likelihood parameters:

```{r, eval=TRUE,echo=TRUE}
mean <- 55.326901
sd <- 16.212602
shape <- (mean^2)/sd
scale <- sd/mean
```

To simulate a new dataset:

```{r, eval=TRUE,echo=TRUE}
z <- rgamma(n=71,shape=shape, scale=scale)
z <- data.frame(1:71,z)
names(z) <- list("ID","myVar")
z <- z[z$myVar>0,]
str(z)
summary(z$myVar)
```

Histogram:

```{r, eval=TRUE,echo=TRUE}
p1 <- ggplot(data=z, aes(x=myVar, y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2)
print(p1)
```

With probability density curve:

```{r, eval=TRUE,echo=TRUE}
stat4 <- stat_function(aes(x = xval, y = ..y..), fun = dgamma, colour="brown", n = length(z$myVar), args = list(shape=shape, rate=(1/scale)))
p1 + stat4
```

Original histogram:

```{r, eval=TRUE,echo=TRUE}
z <- read.csv("kbsampledata.csv", header=TRUE,sep=",")
names(z) <- list("ID","myVar")
z <- z[z$myVar>0,]
str(z)
summary(z)

#histogram
p1 <- ggplot(data=z, aes(x=myVar, y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2)
print(p1)
```

With probability density curve:

```{r, eval=TRUE,echo=TRUE}
gammaPars <- fitdistr(z$myVar,"gamma")
shapeML <- gammaPars$estimate["shape"]
rateML <- gammaPars$estimate["rate"]

stat4 <- stat_function(aes(x = xval, y = ..y..), fun = dgamma, colour="brown", n = length(z$myVar), args = list(shape=shapeML, rate=rateML))
p1 + stat4
```

The simulation performed alright with this data set. The only thing that stuck out to me was the fact that with my original data, the gamma distribution didn't follow the highest peaks of the histogram. Instead, it seemed to follow a smaller value (probably the mean). 
