---
title: "Homework 6"
author: "Emma Privett"
date: "2025-02-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

My data set looks at the length to mass ratio of yellow perch caught in Lake Champlain. To create this data set, I first pulled real values from 71 fish from my experimental data. 
```{r, eval=TRUE,echo=TRUE}
kbdata <- read.csv(file="kbsampledata.csv",header=TRUE, sep = ",")
kblength <- kbdata$length
kbmass <- kbdata$mass
```
From there, I ran a linear regression on the data to generate an r-squared value.
```{r, eval=TRUE,echo=TRUE}
lin_reg <- lm(kbmass~kblength)
summary(lin_reg)
plot(kbdata)
```
I saved the standard deviation and mean from both the length and mass data to apply to a random normal distribution with a sample size of 71.

```{r, eval=TRUE,echo=TRUE}
length_mean<- mean(kblength)
mass_mean <- mean(kbmass)
length_sd <- sd(kblength)
mass_sd <- sd(kbmass)
kblength_rnorm <- rnorm(71, mean=length_mean, sd=length_sd)
kbmass_rnorm <- rnorm(71, mean=mass_mean, sd=mass_sd)
```
To alter this data set, I chose to vary the sample size and compare their r-squared values. I put these new sample sizes in the vector n. These new sample sizes represent the number of fish caught at each site.
```{r, eval=TRUE,echo=TRUE}
n <- c(10,20,50,100)
```
Here is my for loop and the results yielded from it.
```{r, eval=TRUE,echo=TRUE}
for (i in 1:4) {
  new_length<- rnorm(n[i], mean=length_mean, sd=length_sd)
  new_mass <- rnorm(n[i], mean=mass_mean, sd=mass_sd)
  loop_linreg <- lm(new_mass~new_length)
  sumstats<- summary(loop_linreg)
  print(sumstats)
}
```

